---
title: "Forecasting_Fraud"
output: html_document
date: "2026-07-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
#Install it if you haven't.
library(guardianapi)
```

```{r}
#Enter your Guardian API key.
gu_api_key()
```

```{r}
#Find contents with required keywords and period.
fraud <- gu_content(query = '"fraud" AND ("UK" OR "United Kingdom")', from_date = "1999-01-01",
                            to_date = "2024-12-31")
```

```{r}
#Find contents for comparing with forecasts.
fraud1 <- gu_content(query = '"fraud" AND ("UK" OR "United Kingdom")', from_date = "2025-01-01",
                            to_date = "2025-12-31")
```

```{r}
#Inspect the articles obtained.
tibble::glimpse(fraud)
```

# Inspect the contents:

```{r}
table(fraud$type)
```

```{r}
table(fraud$pillar_name)
```

```{r}
table(fraud$production_office)#All English-speaking countries so all contents are in English.
```

```{r}
library(tidyverse)
fraud %>%
  filter(pillar_name == "Sport") %>%
  select(web_title, web_publication_date) %>%
  head(20)
```

```{r}
library(tidyverse)
fraud %>%
  filter(pillar_name == "Arts") %>%
  select(web_title, web_publication_date) %>%
  head(20)
```

```{r}
fraud %>%
  filter(type == "liveblog") %>%
  select(web_title, web_publication_date) %>%
  head(20)
```

```{r}
fraud %>%
  filter(type == "article") %>%
  select(web_title, web_publication_date) %>%
  head(20)
```

```{r}
#Collect only contents with type "article" and pillar name "Lifestyle"/"News"/"Opinion".
fraud %>%
  filter(type == "article" & pillar_name %in% c("Lifestyle","News","Opinion")) %>%
  select(web_title, web_publication_date) %>%
  head(20)
```

# Count the number of articles published each month

```{r}
library(dplyr)
library(lubridate)

monthly_fraud <- fraud %>%
  filter(type == "article",
         pillar_name %in% c("Lifestyle", "News", "Opinion")) %>%
  mutate(month = floor_date(as.Date(web_publication_date), unit = "month")) %>%
  count(month, name = "n_articles")

monthly_fraud

# Save to a csv
write.csv(monthly_fraud,
          file = "fraud_train.csv",
          row.names = FALSE)
```

```{r}
# Also for the tested contents.
monthly_fraud1 <- fraud1 %>%
  filter(type == "article",
         pillar_name %in% c("Lifestyle", "News", "Opinion")) %>%
  mutate(month = floor_date(as.Date(web_publication_date), unit = "month")) %>%
  count(month, name = "n_articles")

monthly_fraud1

# Save to a csv
write.csv(monthly_fraud1,
          file = "fraud_test.csv",
          row.names = FALSE)
```

# Import fraud and macroeconomic data (treat n_articles as fraud)

```{r}
# Import the training dataset
train <- read.csv("fraud&macroeconomic_train.csv")
View(train)

# Import the testing dataset
test <- read.csv("fraud&macroeconomic_test.csv")
View(test)

#The volume of content fetched by the Guardian API may vary across time, so please subject to these two datasets in this project.
```

```{r}
# Name each time series
fraud <- train$fraud
inflation <- train$CPIH
unemployment <- train$unemployment
interest <- train$interest_rate
production <- train$index_of_production_and_industry_sectors
confidence <- train$composite_consumer_confidence
```

```{r}
# Name variables in the test dataset
fraud_test <- test$fraud
inflation_test <- test$CPIH
unemployment_test <- test$unemployment
interest_test <- test$interest_rate
production_test <- test$index_of_production_and_industry_sectors
confidence_test <- test$composite_consumer_confidence
```

# Cross-correlation analysis to find the lead-lag relationships between variables

```{r}
# Test the stationarity of all time series.
library(tseries)

adf.test(fraud)
adf.test(inflation)
adf.test(unemployment)
adf.test(interest)
adf.test(production)
adf.test(confidence)
```

```{r}
# Take first order differencing to check if non-stationary time series become stationary.
adf.test(diff(unemployment))
adf.test(diff(interest))
adf.test(diff(production))
```

```{r}
#Print cross-correlation functions
ccf(fraud, inflation)
ccf(fraud, diff(unemployment))
ccf(fraud, diff(interest))
ccf(fraud, diff(production))
ccf(fraud, confidence)
#Print cross-correlation values
ccf(fraud, inflation, plot = FALSE)
ccf(fraud, diff(unemployment), plot = FALSE)
ccf(fraud, diff(interest), plot = FALSE)
ccf(fraud, diff(production), plot = FALSE)
ccf(fraud, confidence, plot = FALSE)
```

```{r}
ccf(inflation, diff(unemployment))
ccf(inflation, diff(interest))
ccf(inflation, diff(production))
ccf(inflation, confidence)
ccf(inflation, diff(unemployment), plot = FALSE)
ccf(inflation, diff(interest), plot = FALSE)
ccf(inflation, diff(production), plot = FALSE)
ccf(inflation, confidence, plot = FALSE)
```

```{r}
ccf(diff(unemployment), diff(interest))
ccf(diff(unemployment), diff(production))
ccf(diff(unemployment), confidence)
ccf(diff(unemployment), diff(interest), plot = FALSE)
ccf(diff(unemployment), diff(production), plot = FALSE)
ccf(diff(unemployment), confidence, plot = FALSE)
```

```{r}
ccf(diff(interest), diff(production))
ccf(diff(interest), confidence)
ccf(diff(interest), diff(production), plot = FALSE)
ccf(diff(interest), confidence, plot = FALSE)
```

```{r}
ccf(diff(production), confidence)
ccf(diff(production), confidence, plot = FALSE)
```

# Simple linear regression models to understand the linear relationships between variables

```{r}
summary(lm(fraud ~ inflation))
summary(lm(fraud ~ unemployment))
summary(lm(fraud ~ interest))
summary(lm(fraud ~ production))
summary(lm(fraud ~ confidence))
```

```{r}
summary(lm(inflation ~ unemployment))
summary(lm(inflation ~ interest))
summary(lm(inflation ~ production))
summary(lm(inflation ~ confidence))
```

```{r}
summary(lm(unemployment ~ interest))
summary(lm(unemployment ~ production))
summary(lm(unemployment ~ confidence))
```

```{r}
summary(lm(interest ~ production))
summary(lm(interest ~ confidence))
```

```{r}
summary(lm(production ~ confidence))
```

# Build generalised additive models

```{r}
library(mgcv)
gam1 <- gam(fraud ~ s(inflation) + s(unemployment) + s(interest) + s(production) + s(confidence))
summary(gam1)
```

```{r}
gam2 <- gam(fraud ~ s(inflation) + s(unemployment) + s(confidence))
summary(gam2)
```


```{r}
gam3 <- gam(fraud ~ s(unemployment) + s(interest) + s(production) + s(confidence))
summary(gam3)
```

```{r}
# Compare gam1 and gam3
AIC(gam1, gam3)
BIC(gam1, gam3)
anova(gam1, gam3, test = "F")
```

```{r}
# Consider tensor product interactions
gam4 <- gam(fraud ~ s(unemployment) + s(interest) + s(production) + s(confidence) + ti(unemployment, production) + ti(production, confidence) + ti(inflation, confidence))
summary(gam4)
```

```{r}
gam5 <- gam(fraud ~ unemployment + s(interest) + s(production) + confidence + ti(unemployment, production) + ti(production, confidence) + ti(inflation, confidence))
summary(gam5)
```

```{r}
gam6 <- gam(fraud ~ unemployment + s(interest) + s(production) + confidence + ti(unemployment, production) + ti(inflation, confidence))
summary(gam6)
```


```{r}
AIC(gam5, gam6)
BIC(gam5, gam6)
anova(gam5, gam6, test = "F")
```

```{r}
# Consider using the linear relationship for production
gam7 <- gam(fraud ~ unemployment + s(interest) + production + confidence + ti(unemployment, production) + ti(inflation, confidence))
summary(gam7)
```

```{r}
gam8 <- gam(fraud ~s(interest) + production + confidence + ti(unemployment, production) + ti(inflation, confidence))
summary(gam8)
```

```{r}
AIC(gam8)
BIC(gam8)
```

# Build Vector Autoregressive Moving Average models

```{r}
# Build VARMA time series
varma_data0 <- cbind(
  fraud,
  inflation,
  unemployment,
  interest,
  production,
  confidence
)

library(tseries)

varma_ts0 <- ts(varma_data0, start = c(1999, 1), frequency = 12)#Starting from the 1st period of 1999, and there are 12 such periods in a year, i.e. 12 months.
plot(varma_ts0)
```

```{r}
#"Confidence" shows non-stationary trend, thus apply differencing.
adf.test(diff(confidence))
```


```{r}
varma_data1 <- cbind(
  fraud[-1], #Since 1st-order differencing removes the first observation, repeat it for other time series.
  inflation[-1],
  diff(unemployment),
  diff(interest),
  diff(production),
  diff(confidence)
)

varma_ts1 <- ts(varma_data1, start = c(1999, 2), frequency = 12)#Starting from the 2st period of 1999.
plot(varma_ts1)
```

```{r}
# Determine the autoregressive order p, using values from 1 to 30.
library(vars)

VARselect(varma_data1, lag.max = 30, type = "const")
#p = 1 or 4 is suggested
```

```{r}
install.packages("MTS")
library(MTS)
```

```{r}
# Fit VARMA.
full10 <- VARMA(varma_data1, p = 1, q = 0)
full11 <- VARMA(varma_data1, p = 1, q = 1)
full12 <- VARMA(varma_data1, p = 1, q = 2)
full13 <- VARMA(varma_data1, p = 1, q = 3)
```

```{r}
full40 <- VARMA(varma_data1, p = 4, q = 0)
full41 <- VARMA(varma_data1, p = 4, q = 1)
full42 <- VARMA(varma_data1, p = 4, q = 2)
full43 <- VARMA(varma_data1, p = 4, q = 3)
```

```{r}
c(full10$aic, full10$bic)
c(full11$aic, full11$bic)
c(full12$aic, full12$bic)
c(full13$aic, full13$bic)
c(full40$aic, full40$bic)
c(full41$aic, full41$bic)
c(full42$aic, full42$bic)
c(full43$aic, full43$bic)
# Select full10 and full 40.
```

```{r}
# Use only macroeconomic indicators with significant cross-correlations with fraud.
varma_data2 <- cbind(
  fraud[-1], #Since 1st-order differencing removes the first observation, repeat it for other time series.
  inflation[-1],
  diff(unemployment),
  diff(confidence)
)
```

```{r}
VARselect(varma_data2, lag.max = 30, type = "const")
```

```{r}
part10 <- VARMA(varma_data2, p = 1, q = 0)
part11 <- VARMA(varma_data2, p = 1, q = 1)
part12 <- VARMA(varma_data2, p = 1, q = 2)
part13 <- VARMA(varma_data2, p = 1, q = 3)
```

```{r}
part20 <- VARMA(varma_data2, p = 2, q = 0)
part21 <- VARMA(varma_data2, p = 2, q = 1)
part22 <- VARMA(varma_data2, p = 2, q = 2)
part23 <- VARMA(varma_data2, p = 2, q = 3)
```

```{r}
part40 <- VARMA(varma_data2, p = 4, q = 0)
part41 <- VARMA(varma_data2, p = 4, q = 1)
part42 <- VARMA(varma_data2, p = 4, q = 2)
part43 <- VARMA(varma_data2, p = 4, q = 3)
```

```{r}
c(part10$aic, part10$bic)
c(part11$aic, part11$bic)
c(part12$aic, part12$bic)
c(part13$aic, part13$bic)
c(part20$aic, part20$bic)
c(part21$aic, part21$bic)
c(part22$aic, part22$bic)
c(part23$aic, part23$bic)
c(part40$aic, part40$bic)
c(part41$aic, part41$bic)
c(part42$aic, part42$bic)
c(part43$aic, part43$bic)
#Select part40 and part11.
```

```{r}
# Forecasting
pred1 <- VARMApred(
  full10,
  h = 12
)
```

```{r}
# Convert differences to their original values, starting from the last value of the training dataset.
pred1$pred[, 3] <- unemployment[length(unemployment)] + cumsum(pred1$pred[, 3])
pred1$pred[, 4] <- interest[length(interest)] + cumsum(pred1$pred[, 4])
pred1$pred[, 5] <- production[length(production)] + cumsum(pred1$pred[, 5])
pred1$pred[, 6] <- confidence[length(confidence)] + cumsum(pred1$pred[, 6])
pred1$pred[, 3]
pred1$pred[, 4]
pred1$pred[, 5]
pred1$pred[, 6]
```

```{r}
install.packages("forecast")
library(forecast)
```

Note that $\mathrm{MASE} = \frac{\frac{1}{h}\displaystyle\sum_{t=1}^{h}\left|y_t-\hat{y}_t\right|}{\frac{1}{n-1}\displaystyle\sum_{t=2}^{n}\left|y_t-y_{t-1}\right|}$.

```{r}
# Forecasting accuracy assessment
accuracy(object = pred1$pred[, 1], x = fraud_test)
errors <- fraud_test - pred1$pred[, 1]
MAE <- mean(abs(errors))
scale <- mean(abs(diff(fraud)))
MASE <- MAE / scale
MASE
ACF1 <- Acf(errors, plot = FALSE)$acf[2]
ACF1

accuracy(object = pred1$pred[, 2], x = inflation_test)
errors <- inflation_test - pred1$pred[, 2]
MAE <- mean(abs(errors))
scale <- mean(abs(diff(inflation)))
MASE <- MAE / scale
MASE
ACF1 <- Acf(errors, plot = FALSE)$acf[2]
ACF1

accuracy(object = pred1$pred[, 3], x = unemployment_test)
errors <- unemployment_test - pred1$pred[, 3]
MAE <- mean(abs(errors))
scale <- mean(abs(diff(unemployment)))
MASE <- MAE / scale
MASE
ACF1 <- Acf(errors, plot = FALSE)$acf[2]
ACF1

accuracy(object = pred1$pred[, 4], x = interest_test)
errors <- interest_test - pred1$pred[, 4]
MAE <- mean(abs(errors))
scale <- mean(abs(diff(interest)))
MASE <- MAE / scale
MASE
ACF1 <- Acf(errors, plot = FALSE)$acf[2]
ACF1

accuracy(object = pred1$pred[, 5], x = production_test)
errors <- production_test - pred1$pred[, 5]
MAE <- mean(abs(errors))
scale <- mean(abs(diff(production)))
MASE <- MAE / scale
MASE
ACF1 <- Acf(errors, plot = FALSE)$acf[2]
ACF1

accuracy(object = pred1$pred[, 6], x = confidence_test)
errors <- confidence_test - pred1$pred[, 6]
MAE <- mean(abs(errors))
scale <- mean(abs(diff(confidence)))
MASE <- MAE / scale
MASE
ACF1 <- Acf(errors, plot = FALSE)$acf[2]
ACF1
```

```{r}
pred2 <- VARMApred(
  full40,
  h = 12
)
```

```{r}
pred2$pred[, 3] <- unemployment[length(unemployment)] + cumsum(pred2$pred[, 3])
pred2$pred[, 4] <- interest[length(interest)] + cumsum(pred2$pred[, 4])
pred2$pred[, 5] <- production[length(production)] + cumsum(pred2$pred[, 5])
pred2$pred[, 6] <- confidence[length(confidence)] + cumsum(pred2$pred[, 6])
pred2$pred[, 3]
pred2$pred[, 4]
pred2$pred[, 5]
pred2$pred[, 6]
```

```{r}
accuracy(object = pred2$pred[, 1], x = fraud_test)
errors <- fraud_test - pred2$pred[, 1]
MAE <- mean(abs(errors))
scale <- mean(abs(diff(fraud)))
MASE <- MAE / scale
MASE
ACF1 <- Acf(errors, plot = FALSE)$acf[2]
ACF1

accuracy(object = pred2$pred[, 2], x = inflation_test)
errors <- inflation_test - pred2$pred[, 2]
MAE <- mean(abs(errors))
scale <- mean(abs(diff(inflation)))
MASE <- MAE / scale
MASE
ACF1 <- Acf(errors, plot = FALSE)$acf[2]
ACF1

accuracy(object = pred2$pred[, 3], x = unemployment_test)
errors <- unemployment_test - pred2$pred[, 3]
MAE <- mean(abs(errors))
scale <- mean(abs(diff(unemployment)))
MASE <- MAE / scale
MASE
ACF1 <- Acf(errors, plot = FALSE)$acf[2]
ACF1

accuracy(object = pred2$pred[, 4], x = interest_test)
errors <- interest_test - pred2$pred[, 4]
MAE <- mean(abs(errors))
scale <- mean(abs(diff(interest)))
MASE <- MAE / scale
MASE
ACF1 <- Acf(errors, plot = FALSE)$acf[2]
ACF1

accuracy(object = pred2$pred[, 5], x = production_test)
errors <- production_test - pred2$pred[, 5]
MAE <- mean(abs(errors))
scale <- mean(abs(diff(production)))
MASE <- MAE / scale
MASE
ACF1 <- Acf(errors, plot = FALSE)$acf[2]
ACF1

accuracy(object = pred2$pred[, 6], x = confidence_test)
errors <- confidence_test - pred2$pred[, 6]
MAE <- mean(abs(errors))
scale <- mean(abs(diff(confidence)))
MASE <- MAE / scale
MASE
ACF1 <- Acf(errors, plot = FALSE)$acf[2]
ACF1
```

```{r}
pred3 <- VARMApred(
  part40,
  h = 12
)
```

```{r}
pred3$pred[, 3] <- unemployment[length(unemployment)] + cumsum(pred3$pred[, 3])
pred3$pred[, 4] <- confidence[length(confidence)] + cumsum(pred3$pred[, 4])
pred3$pred[, 3]
pred3$pred[, 4]
```

```{r}
accuracy(object = pred3$pred[, 1], x = fraud_test)
errors <- fraud_test - pred3$pred[, 1]
MAE <- mean(abs(errors))
scale <- mean(abs(diff(fraud)))
MASE <- MAE / scale
MASE
ACF1 <- Acf(errors, plot = FALSE)$acf[2]
ACF1

accuracy(object = pred3$pred[, 2], x = inflation_test)
errors <- inflation_test - pred3$pred[, 2]
MAE <- mean(abs(errors))
scale <- mean(abs(diff(inflation)))
MASE <- MAE / scale
MASE
ACF1 <- Acf(errors, plot = FALSE)$acf[2]
ACF1

accuracy(object = pred3$pred[, 3], x = unemployment_test)
errors <- unemployment_test - pred3$pred[, 3]
MAE <- mean(abs(errors))
scale <- mean(abs(diff(unemployment)))
MASE <- MAE / scale
MASE
ACF1 <- Acf(errors, plot = FALSE)$acf[2]
ACF1

accuracy(object = pred3$pred[, 4], x = confidence_test)
errors <- confidence_test - pred3$pred[, 4]
MAE <- mean(abs(errors))
scale <- mean(abs(diff(confidence)))
MASE <- MAE / scale
MASE
ACF1 <- Acf(errors, plot = FALSE)$acf[2]
ACF1
```

```{r}
pred4 <- VARMApred(
  part11,
  h = 12
)
```

```{r}
pred4$pred[, 3] <- unemployment[length(unemployment)] + cumsum(pred4$pred[, 3])
pred4$pred[, 4] <- confidence[length(confidence)] + cumsum(pred4$pred[, 4])
pred4$pred[, 3]
pred4$pred[, 4]
```

```{r}
accuracy(object = pred4$pred[, 1], x = fraud_test)
errors <- fraud_test - pred4$pred[, 1]
MAE <- mean(abs(errors))
scale <- mean(abs(diff(fraud)))
MASE <- MAE / scale
MASE
ACF1 <- Acf(errors, plot = FALSE)$acf[2]
ACF1

accuracy(object = pred4$pred[, 2], x = inflation_test)
errors <- inflation_test - pred4$pred[, 2]
MAE <- mean(abs(errors))
scale <- mean(abs(diff(inflation)))
MASE <- MAE / scale
MASE
ACF1 <- Acf(errors, plot = FALSE)$acf[2]
ACF1

accuracy(object = pred4$pred[, 3], x = unemployment_test)
errors <- unemployment_test - pred4$pred[, 3]
MAE <- mean(abs(errors))
scale <- mean(abs(diff(unemployment)))
MASE <- MAE / scale
MASE
ACF1 <- Acf(errors, plot = FALSE)$acf[2]
ACF1

accuracy(object = pred4$pred[, 4], x = confidence_test)
errors <- confidence_test - pred4$pred[, 4]
MAE <- mean(abs(errors))
scale <- mean(abs(diff(confidence)))
MASE <- MAE / scale
MASE
ACF1 <- Acf(errors, plot = FALSE)$acf[2]
ACF1
```

# Build the Extreme Gradient Boosting model

```{r}
install.packages("xgboost")
library(xgboost)
```

```{r}
# Prepare the training dataset 
train$month <- as.Date(train$month) #Treat months as date variables

library(dplyr)

train_data <- train %>%
  arrange(month) %>%
  mutate(
    inflation_lag1 = lag(inflation, 1), # inflation means CPIH
    inflation_lag4 = lag(inflation, 4),
    unemployment_lag1 = lag(unemployment, 1),
    unemployment_lag21 = lag(unemployment, 21),
    interest_lag1 = lag(interest, 1), # interest means interest_rate
    production_lag1 = lag(production, 1), # production means index_of_production_and_industry_sectors
    confidence_lag1 = lag(confidence, 1) # confidence means composite_consumer_confidence
  ) %>%
  na.omit()      #Clean all rows containing NAs produced by lagging.

head(train_data) #Inspect the training dataset
```

```{r}
# Select the predictor columns
predictor_names <- c(
  "CPIH",
  "unemployment",
  "interest_rate",
  "index_of_production_and_industry_sectors",
  "composite_consumer_confidence",
  "inflation_lag1",
  "inflation_lag4",
  "unemployment_lag1",
  "unemployment_lag21",
  "interest_lag1",
  "production_lag1",
  "confidence_lag1"
)
```

```{r}
# Create the predictor matrix and the response variable
X_train <- as.matrix(train_data[, predictor_names])
y_train <- train_data$fraud
```

```{r}
# Construct xgb.DMatrix object
dtrain <- xgb.DMatrix(
  data = X_train,
  label = y_train
)
```

```{r}
# Train the model
xgb_model <- xgb.train(
  params = xgb.params(objective = "reg:squarederror"), #reggression objective with squared loss
  data = dtrain,
  nrounds = 1000, #maximum of 1000 boosting iterations
  verbose = 1 #it will print information about performance
)
```

# Forecasting

```{r}
# Investigate whether there are seasonal patterns in the macroeconomic indicators that will be fitted in a univariate AR(I)MA or SAR(I)MA model.
acf(inflation, lag.max = 50)
acf(interest, lag.max = 50)
```

```{r}
install.packages("forecast")
library(forecast)
```

```{r}
# Determine the difference order, d
ndiffs(inflation)
ndiffs(interest)
```

```{r}
# PACF for determining p, ACF for q, of ARIMA
acf(diff(inflation))
pacf(diff(inflation))
acf(diff(interest))
pacf(diff(interest))
```

```{r}
fit11 <- arima(inflation, order = c(3,1,0))
fit12 <- arima(inflation, order = c(3,1,1))
fit13 <- arima(inflation, order = c(3,1,2))
fit14 <- arima(inflation, order = c(4,1,0))
fit15 <- arima(inflation, order = c(4,1,1))
fit16 <- arima(inflation, order = c(4,1,2))
AIC(fit11, fit12, fit13, fit14, fit15, fit16)
BIC(fit11, fit12, fit13, fit14, fit15, fit16)#fit13 is used
```

```{r}
fit21 <- arima(interest, order = c(2,1,1))
fit22 <- arima(interest, order = c(3,1,1))
AIC(fit21, fit22)#It suggests fit21, but the difference is very small.
BIC(fit21, fit22)#It suggests fit22, and the difference is more significant. Also there is a spike at lag 3, so fit22 is used.
```

```{r}
pred_inflation <- predict(fit13, n.ahead = 12)
pred_inflation$pred

pred_interest <- predict(fit22, n.ahead = 12)
pred_interest$pred
```

```{r}
#Build the VARMA model for the other 3 macroeconomic indicators.
varma_data2 <- cbind(
  diff(unemployment),
  diff(production),
  diff(confidence)
)
```

```{r}
library(vars)

VARselect(varma_data2, lag.max = 30, type = "const")#Try p = 1,2,8
```

```{r}
fit30 <- VARMA(varma_data2, p = 1, q = 0)
fit31 <- VARMA(varma_data2, p = 1, q = 1)
fit32 <- VARMA(varma_data2, p = 1, q = 2)
#fit34 <- VARMA(varma_data2, p = 1, q = 3) #fail to fit
```

```{r}
fit40 <- VARMA(varma_data2, p = 2, q = 0)
fit41 <- VARMA(varma_data2, p = 2, q = 1)
fit42 <- VARMA(varma_data2, p = 2, q = 2)
fit43 <- VARMA(varma_data2, p = 2, q = 3)
```

```{r}
fit50 <- VARMA(varma_data2, p = 8, q = 0)
fit51 <- VARMA(varma_data2, p = 8, q = 1)
fit52 <- VARMA(varma_data2, p = 8, q = 2)
fit53 <- VARMA(varma_data2, p = 8, q = 3)
```

```{r}
c(fit30$aic, fit30$bic)
c(fit31$aic, fit31$bic)
c(fit32$aic, fit32$bic)
c(fit40$aic, fit40$bic)
c(fit41$aic, fit41$bic)
c(fit42$aic, fit42$bic)
c(fit43$aic, fit43$bic)
c(fit50$aic, fit50$bic)
c(fit51$aic, fit51$bic)
c(fit52$aic, fit52$bic)
c(fit53$aic, fit53$bic) #fit31 and fit41 are chosen
```

```{r}
# Forecasting
pred5 <- VARMApred(
  fit31,
  h = 12
)
```

```{r}
pred5$pred[, 1] <- unemployment[length(unemployment)] + cumsum(pred5$pred[, 1])
pred5$pred[, 2] <- production[length(production)] + cumsum(pred5$pred[, 2])
pred5$pred[, 3] <- confidence[length(confidence)] + cumsum(pred5$pred[, 3])
pred5$pred[, 1]
pred5$pred[, 2]
pred5$pred[, 3]
```

```{r}
accuracy(object = pred5$pred[, 1], x = unemployment_test)
errors <- unemployment_test - pred5$pred[, 1]
MAE <- mean(abs(errors))
scale <- mean(abs(diff(unemployment)))
MASE <- MAE / scale
MASE
ACF1 <- Acf(errors, plot = FALSE)$acf[2]
ACF1

accuracy(object = pred5$pred[, 2], x = production_test)
errors <- production_test - pred5$pred[, 2]
MAE <- mean(abs(errors))
scale <- mean(abs(diff(production)))
MASE <- MAE / scale
MASE
ACF1 <- Acf(errors, plot = FALSE)$acf[2]
ACF1

accuracy(object = pred5$pred[, 3], x = confidence_test)
errors <- confidence_test - pred5$pred[, 3]
MAE <- mean(abs(errors))
scale <- mean(abs(diff(confidence)))
MASE <- MAE / scale
MASE
ACF1 <- Acf(errors, plot = FALSE)$acf[2]
ACF1
```

```{r}
pred6 <- VARMApred(
  fit41,
  h = 12
)
```

```{r}
pred6$pred[, 1] <- unemployment[length(unemployment)] + cumsum(pred6$pred[, 1])
pred6$pred[, 2] <- production[length(production)] + cumsum(pred6$pred[, 2])
pred6$pred[, 3] <- confidence[length(confidence)] + cumsum(pred6$pred[, 3])
pred6$pred[, 1]
pred6$pred[, 2]
pred6$pred[, 3]
```

```{r}
accuracy(object = pred6$pred[, 1], x = unemployment_test)
errors <- unemployment_test - pred6$pred[, 1]
MAE <- mean(abs(errors))
scale <- mean(abs(diff(unemployment)))
MASE <- MAE / scale
MASE
ACF1 <- Acf(errors, plot = FALSE)$acf[2]
ACF1

accuracy(object = pred6$pred[, 2], x = production_test)
errors <- production_test - pred6$pred[, 2]
MAE <- mean(abs(errors))
scale <- mean(abs(diff(production)))
MASE <- MAE / scale
MASE
ACF1 <- Acf(errors, plot = FALSE)$acf[2]
ACF1

accuracy(object = pred6$pred[, 3], x = confidence_test)
errors <- confidence_test - pred6$pred[, 3]
MAE <- mean(abs(errors))
scale <- mean(abs(diff(confidence)))
MASE <- MAE / scale
MASE
ACF1 <- Acf(errors, plot = FALSE)$acf[2]
ACF1
#fit41 (VARMA(2, 1)) is preferred than fit31 (VARMA(1, 1))
```

```{r}
#Prepare the test data for GAM and predict fraud.
future_data_gam <- data.frame(
  inflation = c(3.517067,3.555545,3.606053,3.657496,3.701093,3.731660,3.747602,3.750231,3.742752,3.729203,3.713542,3.699003),
  unemployment = c(4.425784,4.444627,4.458527,4.470378,4.480064,4.488074,4.494556,4.499670,4.503557,4.506347,4.508163,4.509115),
  interest = c(4.443286,4.475055,4.484013,4.481471,4.483958,4.486428,4.486176,4.486018,4.486403,4.486491,4.486415,4.486438),
  production = c(98.15264,98.16309,98.15955,98.14028,98.11178,98.07649,98.03693,97.99552,97.95404,97.91370,97.87530,97.83932),
  confidence = c(-12.17586,-12.12319,-12.08893,-12.03765,-11.97719,-11.91549,-11.85512,-11.79842,-11.74654,-11.70011,-11.65936,-11.62426)
)

fraud_forecast_gam <- predict(gam8, newdata = future_data_gam)
fraud_forecast_gam
```

```{r}
#Prepare the test data for XGBoost and predict fraud.
future_data_xgb <- data.frame(
  CPIH = c(3.517067,3.555545,3.606053,3.657496,3.701093,3.731660,3.747602,3.750231,3.742752,3.729203,3.713542,3.699003),
  unemployment = c(4.425784,4.444627,4.458527,4.470378,4.480064,4.488074,4.494556,4.499670,4.503557,4.506347,4.508163,4.509115),
  interest_rate = c(4.443286,4.475055,4.484013,4.481471,4.483958,4.486428,4.486176,4.486018,4.486403,4.486491,4.486415,4.486438),
  index_of_production_and_industry_sectors = c(98.15264,98.16309,98.15955,98.14028,98.11178,98.07649,98.03693,97.99552,97.95404,97.91370,97.87530,97.83932),
  composite_consumer_confidence = c(-12.17586,-12.12319,-12.08893,-12.03765,-11.97719,-11.91549,-11.85512,-11.79842,-11.74654,-11.70011,-11.65936,-11.62426),
  inflation_lag1 = c(3.5,3.517067,3.555545,3.606053,3.657496,3.701093,3.731660,3.747602,3.750231,3.742752,3.729203,3.713542),
  inflation_lag4 = c(2.6,3.2,3.5,3.5,3.517067,3.555545,3.606053,3.657496,3.701093,3.731660,3.747602,3.750231),
  unemployment_lag1 = c(4.4,4.425784,4.444627,4.458527,4.470378,4.480064,4.488074,4.494556,4.499670,4.503557,4.506347,4.508163),
  unemployment_lag21 = c(4.0,4.2,4.3,4.3,4.1,4.0,4.0,3.9,4.1,4.2,4.3,4.4),
  interest_lag1 = c(4.4345,4.443286,4.475055,4.484013,4.481471,4.483958,4.486428,4.486176,4.486018,4.486403,4.486491,4.486415),
  production_lag1 = c(97.9496,98.15264,98.16309,98.15955,98.14028,98.11178,98.07649,98.03693,97.99552,97.95404,97.91370,97.87530),
  confidence_lag1 = c(-12.00,-12.17586,-12.12319,-12.08893,-12.03765,-11.97719,-11.91549,-11.85512,-11.79842,-11.74654,-11.70011,-11.65936)
)

future_matrix <- xgb.DMatrix(as.matrix(future_data_xgb))

fraud_forecast_xgb <- predict(xgb_model,
                          future_matrix)
fraud_forecast_xgb
```

# Permutation feature importance of the XGBoost model

```{r}
install.packages("iml")
library(iml)
```

```{r}
predictor <- Predictor$new(
    model = xgb_model,
    data = future_data_xgb,
    y = fraud_test,
    predict.function = function(model, newdata)
        predict(model, newdata = xgb.DMatrix(as.matrix(newdata)))
)

set.seed(123) #Set the random state so the results are reproducible

perm_imp <- FeatureImp$new(
    predictor,
    loss = "rmse"
)

plot(perm_imp)

perm_imp$results #numerical results
```

# Model evaluation (for GAM and XGBoost)

```{r}
# Accuracy measures of inflation and interest forecasting by ARIMA
library(forecast)

forecast::accuracy(object = as.numeric(pred_inflation$pred), x = inflation_test)
errors <- inflation_test - pred_inflation$pred
MAE <- mean(abs(errors))
scale <- mean(abs(diff(inflation)))
MASE <- MAE / scale
MASE
ACF1 <- Acf(errors, plot = FALSE)$acf[2]
ACF1

forecast::accuracy(object = as.numeric(pred_interest$pred), x = interest_test)
errors <- interest_test - pred_interest$pred
MAE <- mean(abs(errors))
scale <- mean(abs(diff(interest)))
MASE <- MAE / scale
MASE
ACF1 <- Acf(errors, plot = FALSE)$acf[2]
ACF1
```

```{r}
# Accuracy measures of the fraud level forecasts by GAM
forecast::accuracy(object = fraud_forecast_gam, x = fraud_test)
errors <- fraud_test - fraud_forecast_gam
MAE <- mean(abs(errors))
scale <- mean(abs(diff(fraud)))
MASE <- MAE / scale
MASE
ACF1 <- Acf(errors, plot = FALSE)$acf[2]
ACF1
```

```{r}
# Accuracy measures of the fraud level forecasts by XGBoost
forecast::accuracy(object = fraud_forecast_xgb, x = fraud_test)
errors <- fraud_test - fraud_forecast_xgb
MAE <- mean(abs(errors))
scale <- mean(abs(diff(fraud)))
MASE <- MAE / scale
MASE
ACF1 <- Acf(errors, plot = FALSE)$acf[2]
ACF1
```

# Granger causality test

```{r}
#Create data frames.
data_inf <- data.frame(
  fraud,
  inflation
)
data_u <- data.frame(
  fraud = fraud[-1],
  unemployment = diff(unemployment)
)
data_int <- data.frame(
  fraud = fraud[-1],
  interest = diff(interest)
)
data_p <- data.frame(
  fraud = fraud[-1],
  production = diff(production)
)
data_c <- data.frame(
  fraud = fraud[-1],
  confidence = diff(confidence)
)
```

```{r}
library(vars)

VARselect(data_inf, lag.max = 30, type = "const")
VARselect(data_u, lag.max = 30, type = "const")
VARselect(data_int, lag.max = 30, type = "const")
VARselect(data_p, lag.max = 30, type = "const")
VARselect(data_c, lag.max = 30, type = "const")
```

```{r}
causality(vars::VAR(data_inf, p = 1, type = "const"), cause = "inflation")
causality(vars::VAR(data_inf, p = 4, type = "const"), cause = "inflation")
causality(vars::VAR(data_inf, p = 13, type = "const"), cause = "inflation")
causality(vars::VAR(data_inf, p = 25, type = "const"), cause = "inflation")
```

```{r}
causality(vars::VAR(data_u, p = 2, type = "const"), cause = "unemployment")
causality(vars::VAR(data_u, p = 4, type = "const"), cause = "unemployment")
```

```{r}
causality(vars::VAR(data_int, p = 1, type = "const"), cause = "interest")
causality(vars::VAR(data_int, p = 4, type = "const"), cause = "interest")
```

```{r}
causality(vars::VAR(data_p, p = 1, type = "const"), cause = "production")
causality(vars::VAR(data_p, p = 2, type = "const"), cause = "production")
causality(vars::VAR(data_p, p = 4, type = "const"), cause = "production")
```

```{r}
causality(vars::VAR(data_c, p = 1, type = "const"), cause = "confidence")
causality(vars::VAR(data_c, p = 4, type = "const"), cause = "confidence")
```




