---
title: "Forecasting_Fraud"
output: html_document
date: "2026-07-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
#Install it if you haven't.
library(guardianapi)
```

```{r}
#Enter your Guardian API key.
gu_api_key()
```

```{r}
#Find contents with required keywords and period.
fraud <- gu_content(query = '"fraud" AND ("UK" OR "United Kingdom")', from_date = "1999-01-01",
                            to_date = "2024-12-31")
```

```{r}
#Find contents for comparing with forecasts.
fraud1 <- gu_content(query = '"fraud" AND ("UK" OR "United Kingdom")', from_date = "2025-01-01",
                            to_date = "2025-12-31")
```

```{r}
#Inspect the articles obtained.
tibble::glimpse(fraud)
```

# Inspect the contents:

```{r}
table(fraud$type)
```

```{r}
table(fraud$pillar_name)
```

```{r}
table(fraud$production_office)#All English-speaking countries so all contents are in English.
```

```{r}
library(tidyverse)
fraud %>%
  filter(pillar_name == "Sport") %>%
  select(web_title, web_publication_date) %>%
  head(20)
```

```{r}
library(tidyverse)
fraud %>%
  filter(pillar_name == "Arts") %>%
  select(web_title, web_publication_date) %>%
  head(20)
```

```{r}
fraud %>%
  filter(type == "liveblog") %>%
  select(web_title, web_publication_date) %>%
  head(20)
```

```{r}
fraud %>%
  filter(type == "article") %>%
  select(web_title, web_publication_date) %>%
  head(20)
```

```{r}
#Collect only contents with type "article" and pillar name "Lifestyle"/"News"/"Opinion".
fraud %>%
  filter(type == "article" & pillar_name %in% c("Lifestyle","News","Opinion")) %>%
  select(web_title, web_publication_date) %>%
  head(20)
```

# Count the number of articles published each month

```{r}
library(dplyr)
library(lubridate)

monthly_fraud <- fraud %>%
  filter(type == "article",
         pillar_name %in% c("Lifestyle", "News", "Opinion")) %>%
  mutate(month = floor_date(as.Date(web_publication_date), unit = "month")) %>%
  count(month, name = "n_articles")

monthly_fraud

# Save to a csv
write.csv(monthly_fraud,
          file = "fraud_train.csv",
          row.names = FALSE)
```

```{r}
# Also for the tested contents.
monthly_fraud1 <- fraud1 %>%
  filter(type == "article",
         pillar_name %in% c("Lifestyle", "News", "Opinion")) %>%
  mutate(month = floor_date(as.Date(web_publication_date), unit = "month")) %>%
  count(month, name = "n_articles")

monthly_fraud1

# Save to a csv
write.csv(monthly_fraud1,
          file = "fraud_test.csv",
          row.names = FALSE)
```

# Import fraud and macroeconomic data (treat n_articles as fraud)

```{r}
# Import the training dataset
train <- read.csv("fraud&macroeconomic_train.csv")
View(train)

# Import the testing dataset
test <- read.csv("fraud&macroeconomic_test.csv")
View(test)
```

```{r}
# Name each time series
fraud <- train$fraud
inflation <- train$CPIH
unemployment <- train$unemployment
interest <- train$interest_rate
production <- train$index_of_production_and_industry_sectors
confidence <- train$composite_consumer_confidence
```

# Cross-correlation analysis to find the lead-lag relationships between variables

```{r}
# Test the stationarity of all time series.
library(tseries)

adf.test(fraud)
adf.test(inflation)
adf.test(unemployment)
adf.test(interest)
adf.test(production)
adf.test(confidence)
```

```{r}
# Take first order differencing to check if non-stationary time series become stationary.
adf.test(diff(unemployment))
adf.test(diff(interest))
adf.test(diff(production))
```

```{r}
#Print cross-correlation functions
ccf(fraud, inflation)
ccf(fraud, diff(unemployment))
ccf(fraud, diff(interest))
ccf(fraud, diff(production))
ccf(fraud, confidence)
#Print cross-correlation values
ccf(fraud, inflation, plot = FALSE)
ccf(fraud, diff(unemployment), plot = FALSE)
ccf(fraud, diff(interest), plot = FALSE)
ccf(fraud, diff(production), plot = FALSE)
ccf(fraud, confidence, plot = FALSE)
```

```{r}
ccf(inflation, diff(unemployment))
ccf(inflation, diff(interest))
ccf(inflation, diff(production))
ccf(inflation, confidence)
ccf(inflation, diff(unemployment), plot = FALSE)
ccf(inflation, diff(interest), plot = FALSE)
ccf(inflation, diff(production), plot = FALSE)
ccf(inflation, confidence, plot = FALSE)
```

```{r}
ccf(diff(unemployment), diff(interest))
ccf(diff(unemployment), diff(production))
ccf(diff(unemployment), confidence)
ccf(diff(unemployment), diff(interest), plot = FALSE)
ccf(diff(unemployment), diff(production), plot = FALSE)
ccf(diff(unemployment), confidence, plot = FALSE)
```

```{r}
ccf(diff(interest), diff(production))
ccf(diff(interest), confidence)
ccf(diff(interest), diff(production), plot = FALSE)
ccf(diff(interest), confidence, plot = FALSE)
```

```{r}
ccf(diff(production), confidence)
ccf(diff(production), confidence, plot = FALSE)
```

# Simple linear regression models to understand the linear relationships between variables

```{r}
summary(lm(fraud ~ inflation))
summary(lm(fraud ~ unemployment))
summary(lm(fraud ~ interest))
summary(lm(fraud ~ production))
summary(lm(fraud ~ confidence))
```

```{r}
summary(lm(inflation ~ unemployment))
summary(lm(inflation ~ interest))
summary(lm(inflation ~ production))
summary(lm(inflation ~ confidence))
```

```{r}
summary(lm(unemployment ~ interest))
summary(lm(unemployment ~ production))
summary(lm(unemployment ~ confidence))
```

```{r}
summary(lm(interest ~ production))
summary(lm(interest ~ confidence))
```

```{r}
summary(lm(production ~ confidence))
```

# Build generalised additive models

```{r}
library(mgcv)
gam1 <- gam(fraud ~ s(inflation) + s(unemployment) + s(interest) + s(production) + s(confidence))
summary(gam1)
```

```{r}
gam2 <- gam(fraud ~ s(inflation) + s(unemployment) + s(confidence))
summary(gam2)
```


```{r}
gam3 <- gam(fraud ~ s(unemployment) + s(interest) + s(production) + s(confidence))
summary(gam3)
```

```{r}
# Compare gam1 and gam3
AIC(gam1, gam3)
BIC(gam1, gam3)
anova(gam1, gam3, test = "F")
```

```{r}
# Consider tensor product interactions
gam4 <- gam(fraud ~ s(unemployment) + s(interest) + s(production) + s(confidence) + ti(unemployment, production) + ti(production, confidence) + ti(inflation, confidence))
summary(gam4)
```

```{r}
gam5 <- gam(fraud ~ unemployment + s(interest) + s(production) + confidence + ti(unemployment, production) + ti(production, confidence) + ti(inflation, confidence))
summary(gam5)
```

```{r}
gam6 <- gam(fraud ~ unemployment + s(interest) + s(production) + confidence + ti(unemployment, production) + ti(inflation, confidence))
summary(gam6)
```


```{r}
AIC(gam5, gam6)
BIC(gam5, gam6)
anova(gam5, gam6, test = "F")
```

```{r}
# Consider using the linear relationship for production
gam7 <- gam(fraud ~ unemployment + s(interest) + production + confidence + ti(unemployment, production) + ti(inflation, confidence))
summary(gam7)
```

```{r}
gam8 <- gam(fraud ~s(interest) + production + confidence + ti(unemployment, production) + ti(inflation, confidence))
summary(gam8)
```

```{r}
AIC(gam8)
BIC(gam8)
```

# Build Vector Autoregressive Moving Average models

```{r}
# Build VARMA time series
varma_data0 <- cbind(
  fraud,
  inflation,
  unemployment,
  interest,
  production,
  confidence
)

library(tseries)

varma_ts0 <- ts(varma_data0, start = c(1999, 1), frequency = 12)#Starting from the 1st period of 1999, and there are 12 such periods in a year, i.e. 12 months.
plot(varma_ts0)
```

```{r}
#"Confidence" shows non-stationary trend, thus apply differencing.
adf.test(diff(confidence))
```


```{r}
varma_data1 <- cbind(
  fraud[-1], #Since 1st-order differencing removes the first observation, repeat it for other time series.
  inflation[-1],
  diff(unemployment),
  diff(interest),
  diff(production),
  diff(confidence)
)

varma_ts1 <- ts(varma_data1, start = c(1999, 2), frequency = 12)#Starting from the 2st period of 1999.
plot(varma_ts1)
```

```{r}
# Determine the autoregressive order p, using values from 1 to 30.
library(vars)

VARselect(varma_data1, lag.max = 30, type = "const")
#p = 1 or 4 is suggested
```

```{r}
install.packages("MTS")
library(MTS)
```

```{r}
# Fit VARMA.
full10 <- VARMA(varma_data1, p = 1, q = 0)
full11 <- VARMA(varma_data1, p = 1, q = 1)
full12 <- VARMA(varma_data1, p = 1, q = 2)
full13 <- VARMA(varma_data1, p = 1, q = 3)
```

```{r}
full40 <- VARMA(varma_data1, p = 4, q = 0)
full41 <- VARMA(varma_data1, p = 4, q = 1)
full42 <- VARMA(varma_data1, p = 4, q = 2)
full43 <- VARMA(varma_data1, p = 4, q = 3)
```

```{r}
c(full10$aic, full10$bic)
c(full11$aic, full11$bic)
c(full12$aic, full12$bic)
c(full13$aic, full13$bic)
c(full40$aic, full40$bic)
c(full41$aic, full41$bic)
c(full42$aic, full42$bic)
c(full43$aic, full43$bic)
# Select full10 and full 40.
```

```{r}
# Use only macroeconomic indicators with significant cross-correlations with fraud.
varma_data2 <- cbind(
  fraud[-1], #Since 1st-order differencing removes the first observation, repeat it for other time series.
  inflation[-1],
  diff(unemployment),
  diff(confidence)
)
```

```{r}
VARselect(varma_data2, lag.max = 30, type = "const")
```

```{r}
part10 <- VARMA(varma_data2, p = 1, q = 0)
part11 <- VARMA(varma_data2, p = 1, q = 1)
part12 <- VARMA(varma_data2, p = 1, q = 2)
part13 <- VARMA(varma_data2, p = 1, q = 3)
```

```{r}
part20 <- VARMA(varma_data2, p = 2, q = 0)
part21 <- VARMA(varma_data2, p = 2, q = 1)
part22 <- VARMA(varma_data2, p = 2, q = 2)
part23 <- VARMA(varma_data2, p = 2, q = 3)
```

```{r}
part40 <- VARMA(varma_data2, p = 4, q = 0)
part41 <- VARMA(varma_data2, p = 4, q = 1)
part42 <- VARMA(varma_data2, p = 4, q = 2)
part43 <- VARMA(varma_data2, p = 4, q = 3)
```

```{r}
c(part10$aic, part10$bic)
c(part11$aic, part11$bic)
c(part12$aic, part12$bic)
c(part13$aic, part13$bic)
c(part20$aic, part20$bic)
c(part21$aic, part21$bic)
c(part22$aic, part22$bic)
c(part23$aic, part23$bic)
c(part40$aic, part40$bic)
c(part41$aic, part41$bic)
c(part42$aic, part42$bic)
c(part43$aic, part43$bic)
#Select part40 and part11.
```

```{r}
# Name variables in the test dataset
fraud_test <- test$fraud
inflation_test <- test$CPIH
unemployment_test <- test$unemployment
interest_test <- test$interest_rate
production_test <- test$index_of_production_and_industry_sectors
confidence_test <- test$composite_consumer_confidence
```

```{r}
# Forecasting
pred1 <- VARMApred(
  full10,
  h = 12
)
```

```{r}
# Convert differences to their original values, starting from the last value of the training dataset.
pred1$pred[, 3] <- unemployment[length(unemployment)] + cumsum(pred1$pred[, 3])
pred1$pred[, 4] <- interest[length(interest)] + cumsum(pred1$pred[, 4])
pred1$pred[, 5] <- production[length(production)] + cumsum(pred1$pred[, 5])
pred1$pred[, 6] <- confidence[length(confidence)] + cumsum(pred1$pred[, 6])
pred1$pred[, 3]
pred1$pred[, 4]
pred1$pred[, 5]
pred1$pred[, 6]
```

```{r}
install.packages("forecast")
library(forecast)
```

```{r}
# Forecasting accuracy assessment
accuracy(object = pred1$pred[, 1], x = fraud_test)
errors <- fraud_test - pred1$pred[, 1]
MAE <- mean(abs(errors))
scale <- mean(abs(diff(fraud)))
MASE <- MAE / scale
MASE
ACF1 <- Acf(errors, plot = FALSE)$acf[2]
ACF1

accuracy(object = pred1$pred[, 2], x = inflation_test)
errors <- inflation_test - pred1$pred[, 2]
MAE <- mean(abs(errors))
scale <- mean(abs(diff(inflation)))
MASE <- MAE / scale
MASE
ACF1 <- Acf(errors, plot = FALSE)$acf[2]
ACF1

accuracy(object = pred1$pred[, 3], x = unemployment_test)
errors <- unemployment_test - pred1$pred[, 3]
MAE <- mean(abs(errors))
scale <- mean(abs(diff(unemployment)))
MASE <- MAE / scale
MASE
ACF1 <- Acf(errors, plot = FALSE)$acf[2]
ACF1

accuracy(object = pred1$pred[, 4], x = interest_test)
errors <- interest_test - pred1$pred[, 4]
MAE <- mean(abs(errors))
scale <- mean(abs(diff(interest)))
MASE <- MAE / scale
MASE
ACF1 <- Acf(errors, plot = FALSE)$acf[2]
ACF1

accuracy(object = pred1$pred[, 5], x = production_test)
errors <- production_test - pred1$pred[, 5]
MAE <- mean(abs(errors))
scale <- mean(abs(diff(production)))
MASE <- MAE / scale
MASE
ACF1 <- Acf(errors, plot = FALSE)$acf[2]
ACF1

accuracy(object = pred1$pred[, 6], x = confidence_test)
errors <- confidence_test - pred1$pred[, 6]
MAE <- mean(abs(errors))
scale <- mean(abs(diff(confidence)))
MASE <- MAE / scale
MASE
ACF1 <- Acf(errors, plot = FALSE)$acf[2]
ACF1
```

```{r}
pred2 <- VARMApred(
  full40,
  h = 12
)
```

```{r}
pred2$pred[, 3] <- unemployment[length(unemployment)] + cumsum(pred2$pred[, 3])
pred2$pred[, 4] <- interest[length(interest)] + cumsum(pred2$pred[, 4])
pred2$pred[, 5] <- production[length(production)] + cumsum(pred2$pred[, 5])
pred2$pred[, 6] <- confidence[length(confidence)] + cumsum(pred2$pred[, 6])
pred2$pred[, 3]
pred2$pred[, 4]
pred2$pred[, 5]
pred2$pred[, 6]
```

```{r}
accuracy(object = pred2$pred[, 1], x = fraud_test)
errors <- fraud_test - pred2$pred[, 1]
MAE <- mean(abs(errors))
scale <- mean(abs(diff(fraud)))
MASE <- MAE / scale
MASE
ACF1 <- Acf(errors, plot = FALSE)$acf[2]
ACF1

accuracy(object = pred2$pred[, 2], x = inflation_test)
errors <- inflation_test - pred2$pred[, 2]
MAE <- mean(abs(errors))
scale <- mean(abs(diff(inflation)))
MASE <- MAE / scale
MASE
ACF1 <- Acf(errors, plot = FALSE)$acf[2]
ACF1

accuracy(object = pred2$pred[, 3], x = unemployment_test)
errors <- unemployment_test - pred2$pred[, 3]
MAE <- mean(abs(errors))
scale <- mean(abs(diff(unemployment)))
MASE <- MAE / scale
MASE
ACF1 <- Acf(errors, plot = FALSE)$acf[2]
ACF1

accuracy(object = pred2$pred[, 4], x = interest_test)
errors <- interest_test - pred2$pred[, 4]
MAE <- mean(abs(errors))
scale <- mean(abs(diff(interest)))
MASE <- MAE / scale
MASE
ACF1 <- Acf(errors, plot = FALSE)$acf[2]
ACF1

accuracy(object = pred2$pred[, 5], x = production_test)
errors <- production_test - pred2$pred[, 5]
MAE <- mean(abs(errors))
scale <- mean(abs(diff(production)))
MASE <- MAE / scale
MASE
ACF1 <- Acf(errors, plot = FALSE)$acf[2]
ACF1

accuracy(object = pred2$pred[, 6], x = confidence_test)
errors <- confidence_test - pred2$pred[, 6]
MAE <- mean(abs(errors))
scale <- mean(abs(diff(confidence)))
MASE <- MAE / scale
MASE
ACF1 <- Acf(errors, plot = FALSE)$acf[2]
ACF1
```

```{r}
pred3 <- VARMApred(
  part40,
  h = 12
)
```

```{r}
pred3$pred[, 3] <- unemployment[length(unemployment)] + cumsum(pred3$pred[, 3])
pred3$pred[, 4] <- confidence[length(confidence)] + cumsum(pred3$pred[, 4])
pred3$pred[, 3]
pred3$pred[, 4]
```

```{r}
accuracy(object = pred3$pred[, 1], x = fraud_test)
errors <- fraud_test - pred3$pred[, 1]
MAE <- mean(abs(errors))
scale <- mean(abs(diff(fraud)))
MASE <- MAE / scale
MASE
ACF1 <- Acf(errors, plot = FALSE)$acf[2]
ACF1

accuracy(object = pred3$pred[, 2], x = inflation_test)
errors <- inflation_test - pred3$pred[, 2]
MAE <- mean(abs(errors))
scale <- mean(abs(diff(inflation)))
MASE <- MAE / scale
MASE
ACF1 <- Acf(errors, plot = FALSE)$acf[2]
ACF1

accuracy(object = pred3$pred[, 3], x = unemployment_test)
errors <- unemployment_test - pred3$pred[, 3]
MAE <- mean(abs(errors))
scale <- mean(abs(diff(unemployment)))
MASE <- MAE / scale
MASE
ACF1 <- Acf(errors, plot = FALSE)$acf[2]
ACF1

accuracy(object = pred3$pred[, 4], x = confidence_test)
errors <- confidence_test - pred3$pred[, 4]
MAE <- mean(abs(errors))
scale <- mean(abs(diff(confidence)))
MASE <- MAE / scale
MASE
ACF1 <- Acf(errors, plot = FALSE)$acf[2]
ACF1
```

```{r}
pred4 <- VARMApred(
  part11,
  h = 12
)
```

```{r}
pred4$pred[, 3] <- unemployment[length(unemployment)] + cumsum(pred4$pred[, 3])
pred4$pred[, 4] <- confidence[length(confidence)] + cumsum(pred4$pred[, 4])
pred4$pred[, 3]
pred4$pred[, 4]
```

```{r}
accuracy(object = pred4$pred[, 1], x = fraud_test)
errors <- fraud_test - pred4$pred[, 1]
MAE <- mean(abs(errors))
scale <- mean(abs(diff(fraud)))
MASE <- MAE / scale
MASE
ACF1 <- Acf(errors, plot = FALSE)$acf[2]
ACF1

accuracy(object = pred4$pred[, 2], x = inflation_test)
errors <- inflation_test - pred4$pred[, 2]
MAE <- mean(abs(errors))
scale <- mean(abs(diff(inflation)))
MASE <- MAE / scale
MASE
ACF1 <- Acf(errors, plot = FALSE)$acf[2]
ACF1

accuracy(object = pred4$pred[, 3], x = unemployment_test)
errors <- unemployment_test - pred4$pred[, 3]
MAE <- mean(abs(errors))
scale <- mean(abs(diff(unemployment)))
MASE <- MAE / scale
MASE
ACF1 <- Acf(errors, plot = FALSE)$acf[2]
ACF1

accuracy(object = pred4$pred[, 4], x = confidence_test)
errors <- confidence_test - pred4$pred[, 4]
MAE <- mean(abs(errors))
scale <- mean(abs(diff(confidence)))
MASE <- MAE / scale
MASE
ACF1 <- Acf(errors, plot = FALSE)$acf[2]
ACF1
```

# Build the Extreme Gradient Boosting model






